784 Lab 2: Review of Tidyverse and ggplot

### 2023-01-19 Lab 2  Tidyverse Review/Lab2_Tidyverse_Review_Answers.Rmd

#One method for saving plots:
#pdf("Plot1.pdf")
#carsplot
#dev.off()

#Where do these plots go?!

#Unique to ggplot:
#ggsave("Plot1a.pdf") #can edit image sizes as well
```

+ Simple barplot example
```{r, message=FALSE, warning=FALSE}

ggplot(mtcars, aes(x=as.factor(cyl) )) + 
     geom_bar() + 
    labs(title = "Count of Cars By Cylinders", x = "Number of Cylinders", y = "Count")

ggplot(mtcars, aes(x=as.factor(cyl) )) + 
    geom_bar() + 
    labs(title = "Count of Cars By Cylinders (Horizontal)", x = "Number of Cylinders", y = "Count")+
     coord_flip()

#Grouped barplot
#Calculate percentage (and counts) of cyl by gear type
  d1<-pctgroup(mtcars,cyl,gear)
#Graph
ggplot(d1, aes(x=cyl,y=Percentage, fill=factor(gear),label=round(Percentage,digits=2))) + 
      geom_bar(position="dodge", stat="identity")+
      labs(title = "Percentage of Cars By Cylinders and Gear Type", x = "Number of Cylinders", y = "Percentage",fill="Gear")

### 2023-01-19 Lab 2  Tidyverse Review/Lab2_Tidyverse_Review.Rmd
Add files via upload

  #the glimpse() function is quite similar to base R's str function
glimpse(starwars)
 #What if we want to pull a specific piece of data?
snip<-slice(starwars, 10:11)

@@

d1 %>%
  group_by(species) %>%
  summarise(avg = mean(____, na.rm=TRUE)) %>%
  arrange(desc(avg))
#Hutts are the heaviest!

@@

#What if we wanted to create a new variable that combined a characters name with where they were born?
#This is easy using tidyr (another package in tidyverse suite)
starwars %>%
  unite(name_homeworld, ____) %>% 
  select(name_homeworld)

@@

```{r}
library(nycflights13)

# Load data 1
(band_members)

@@

#Unique to ggplot:
#ggsave("Plot1a.pdf") #can edit image sizes as well
```

## Themes
+ As you've hopefully seen ggplot is highly customization. One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g. color palettes, axes labels, etc.) is to use a ggplot() theme that includes bundles of code. This is what we were doing using theme_bw() above.

+ You can also create custom themes yourself!

```{r}
#Here is a theme I use often

plot_theme = theme(text = element_text(size=11),  #set text size to 11
                   panel.grid.minor = element_blank(), #these next two lines set the grid lines to null
                   panel.grid.major = element_blank(),
                   panel.background = element_blank(),#these lines make the background of the plot blank
                   plot.background = element_rect(fill = "transparent",colour = NA),
                   axis.text = element_text(color="black"), #setting axis text to black 
                   axis.line = element_line(colour = "black"),#setting axis lines to black 
                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),#adjusting the angle of text
                   axis.title = element_text(lineheight=1.1), #increasing height of axis title
                   legend.position = "bottom",  #Placing the legend at the bottom of the figure
                   legend.key = element_rect(colour = "transparent",  fill = "transparent") #setting legend to transparent
)

#Using the theme
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  plot_theme #no () after the custom theme call
```
+ As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
```{r}
#bw
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()

#minimal
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_minimal()

#classic
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_classic()

#If you want even more themes you can use the ggthemes library!
#install.packages("ggthemes")
library(ggthemes)

#WSJ
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_wsj()

#538
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_fivethirtyeight()

#The Economist
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_economist()

```
## Facets
+ Using facets lets you create multi-panel plots based off a given variable. There are two different ggplot functions to use to facet.

1. facet_wrap() "wraps" a 1d ribbon of panels into 2d. Essentially, creating a square grid of panels, with rows and columns defined by combinations of two categorical variables (even if some combinations of faceting values do not exist in data). This is most useful when you want to facet by a single variable (or two)

```{r}

carsplot_facet <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_wrap(~ am)
```

2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables.

```{r}
#facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
#facet_grid(variable~.) will return facets equal to the levels of variable distributed vertically.

carsplot_facet_2 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(vs ~ am)

carsplot_facet_3 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(am ~ .)

carsplot_facet_4 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(. ~ am)
```

>>>>

  #the glimpse() function is quite similar to base R's str function
glimpse(starwars)

 #What if we want to pull a specific piece of data?
snip<-slice(starwars, 10:11)

@@

d1 %>%
  group_by(species) %>%
  summarise(avg = mean(____, na.rm=TRUE)) %>%
  arrange(desc(avg))

@@

#What if we wanted to create a new variable that combined a characters name with where they were born?
#This is easy using tidyr (another package in tidyverse suite)
?unite

starwars %>%
  unite(name_homeworld, ____) %>% 
  select(name_homeworld)

@@

```{r}
# Load data 1
(band_members)

@@

#Unique to ggplot:
#ggsave("Plot1a.pdf") #can edit image sizes as well
```
+ Simple barplot example
```{r, message=FALSE, warning=FALSE}

ggplot(mtcars, aes(x=as.factor(cyl) )) + 
     geom_bar() + 
    labs(title = "Count of Cars By Cylinders", x = "Number of Cylinders", y = "Count")

ggplot(mtcars, aes(x=as.factor(cyl) )) + 
    geom_bar() + 
    labs(title = "Count of Cars By Cylinders (Horizontal)", x = "Number of Cylinders", y = "Count")+
     coord_flip()

#Grouped barplot
#Calculate percentage (and counts) of cyl by gear type
  d1<-pctgroup(mtcars,cyl,gear)
#Graph
ggplot(d1, aes(x=cyl,y=Percentage, fill=factor(gear),label=round(Percentage,digits=2))) + 
      geom_bar(position="dodge", stat="identity")+
      labs(title = "Percentage of Cars By Cylinders and Gear Type", x = "Number of Cylinders", y = "Percentage",fill="Gear")
